---
title: 'Simple 3-state Markov model in R'
author: "The DARTH workgroup"
output:
  pdf_document: default
  html_document: default
---

Developed by the Decision Analysis in R for Technologies in Health (DARTH) workgroup:

Fernando Alarid-Escudero, PhD (1) 

Eva A. Enns, MS, PhD (2)	

M.G. Myriam Hunink, MD, PhD (3,4)

Hawre J. Jalal, MD, PhD (5) 

Eline M. Krijkamp, MSc (3)	

Petros Pechlivanoglou, PhD (6,7)

Alan Yang, MSc (7)

In collaboration of: 		

1. Drug Policy Program, Center for Research and Teaching in Economics (CIDE) - CONACyT, 
   Aguascalientes, Mexico
2. University of Minnesota School of Public Health, Minneapolis, MN, USA
3. Erasmus MC, Rotterdam, The Netherlands
4. Harvard T.H. Chan School of Public Health, Boston, USA
5. University of Pittsburgh Graduate School of Public Health, Pittsburgh, PA, USA
6. University of Toronto, Toronto ON, Canada
7. The Hospital for Sick Children, Toronto ON, Canada

Please cite our publications when using this code:
 
- Jalal H, Pechlivanoglou P, Krijkamp E, Alarid-Escudero F, Enns E, Hunink MG. 
An Overview of R in Health Decision Sciences. Med Decis Making. 2017; 37(3): 735-746. 
https://journals.sagepub.com/doi/abs/10.1177/0272989X16686559
 
- Krijkamp EM, Alarid-Escudero F, Enns EA, Jalal HJ, Hunink MGM, Pechlivanoglou P. 
Microsimulation modeling for health decision sciences using R: A tutorial. 
Med Decis Making. 2018;38(3):400–22. 
https://journals.sagepub.com/doi/abs/10.1177/0272989X18754513
 
- Krijkamp EM, Alarid-Escudero F, Enns E, Pechlivanoglou P, Hunink MM, Jalal H. 
A Multidimensional Array Representation of State-Transition Model Dynamics. 
Med Decis Making. Online First https://doi.org/10.1177/0272989X19893973

Copyright 2017, THE HOSPITAL FOR SICK CHILDREN AND THE COLLABORATING INSTITUTIONS. 
All rights reserved in Canada, the United States and worldwide. Copyright, 
trademarks, trade names and any and all associated intellectual property are 
exclusively owned by THE HOSPITAL FOR Sick CHILDREN and the collaborating 
institutions. These materials may be used, reproduced, modified, distributed 
and adapted with proper attribution.

\newpage

```{r}
rm(list = ls())      # clear memory (removes all the variables from the workspace)
```

# 01 Load packages

```{r}
library(survival)
```

# 02 Load functions

```{r}
# no functions required

```

# 03 Input model parameters

```{r}
# Strategy names
v_names_str <- c("Base Case")  

# Number of strategies
n_str <- length(v_names_str)

# Markov model parameters
v_n  <- c("Healthy", "Sick", "Dead")    # state names

n_states  <- length(v_n)                # number of states
n_t  <- 60                              # number of cycles
n_i <- 10000

r_HD <- 0.02                           # rate of death when healthy
r_HS <- 0.05                            # rate of becoming sick when healthy
r_SD <-  0.1                           # rate of death when sick

tmat <- matrix(NA, n_states, n_states, dimnames = list(v_n,v_n))
tmat["Healthy", "Sick"]  <- 1
tmat["Healthy", "Dead"]  <- 2
tmat["Sick"   , "Dead"]  <- 3

n_trans = sum(tmat >0, na.rm= T) 


#Draw the values for all state transitions. We ll check later if they actuyally transitioned
t_HS <- rexp(n_i,r_HS)
t_HD <- rexp(n_i,r_HD)
t_SD <- rexp(n_i,r_SD)
#t_SH <- rexp(n_i,r_SH)

t_H <- cbind(t_HS,t_HD)
t_S <- cbind(t_SD)

  state1 <- apply(t_H,1, which.min)-1
  time1  <- apply(t_H,1, min)
  state2 <- apply(t_S,1,which.min)
  time2  <- apply(t_S,1,min)
  time2 <- ifelse(state1=="Sick",time1+time2,time1)

  
  
time  <- cbind( t_0 = 0, time1,  time2)
state <- cbind(m_0 = 1, state1, state2)
res <- list("t"= time, "st" = state)

```


# 06 Compute and Plot Epidemiological Outcomes
```{r}
plot(survfit(Surv(time[,3], state[,3]==1)~1), xlim = c(0,n_t))

```
